---
title: "Greenspace Demo"
runtime: shiny_prerendered
output: flexdashboard::flex_dashboard
orientation: columns
theme: simplex
---

```{r, context='setup', include=FALSE}
library(flexdashboard)
library(shiny)
library(raster)
library(leaflet)
#library(DT)
library(magrittr)
library(colorspace)
library(rgeos)
library(ggplot2)
#library(shinyjs)
#library(plotly)
library(DiagrammeR)
library(rgdal)
library(sf)
#library(tidyverse)

#root<-"~/Documents/Exeter/Eservices_Data/"; in.root<-"/Volumes/Pocket Sam/Data/"
#in.root<-"/Users/jm622/apps/"

###################################
# Greenspace
####################################
in.root<-""
dir_shape<-paste(in.root,"www/", sep="")

################################################
# Global FUNCTIONS
###############################################
# Return path to raster of ecosystem service map
 # returns variable to be used for mapping
calc_grnsp <- function(alot,relg,pkgd,golf,sports,play,open, mapdata) {
  # save and remove geometry converting to normal df
  g<-mapdata[,21]
  st_geometry(mapdata) <- NULL 
  names(mapdata)<-c("soa","objectid","lsoa11nm","lsoa11nmw","st_lengths",   
                     "Allotments","Bowling_Green","Cemetery","Golf_Course","Sports" ,          
                     "Play_Space", "Playing_Field","Park_Garden","Religious_Grounds","Tennis_Court",     
                    "Open_Access", "All.Ages","Ages.0.18","Ages.19.69","Ages.70.90" )
  
  # sum  greenspace 
  area<-rep(0,length(mapdata$soa))  
  if(alot) area<-area+mapdata[,c("Allotments")]
  if(relg) area<-area+mapdata[,c("Cemetery")]+mapdata[,c("Religious_Grounds")]
  if(pkgd) area<-area+mapdata[,c("Park_Garden")]
  if(golf) area<-area + mapdata[,c("Golf_Course")]
  if(sports) area<-area + mapdata[,c("Sports")] + mapdata[,c("Bowling_Green")] + mapdata[,c("Tennis_Court")]
  if(play) area<-area + mapdata[,c("Play_Space")] + mapdata[,c("Playing_Field")]
  if(open) area<-area + mapdata[,c("Open_Access")]
  
  mapdata$area<-area
  mapdata$area_per_pop<-round(mapdata$area/mapdata$All.Ages,1)
  mapdata<-mapdata[,c(1,17,21,22)] # DO NOT USE @data !!
  # Reset geometry to original
  st_geometry(mapdata)<-st_geometry(g)
  return(mapdata)
}

# Calculate bar chart data for all of Cornwall
calc_chartdata_1<-function(alot,relg,pkgd,golf,sports,play,open,mapdata){
    names(mapdata)<-c("soa","objectid","lsoa11nm","lsoa11nmw","st_lengths",   
                      "Allotments","Bowling_Green","Cemetery","Golf_Course","Sports" ,          
                      "Play_Space", "Playing_Field","Park_Garden","Religious_Grounds","Tennis_Court",     
                      "Open_Access", "All.Ages","Ages.0.18","Ages.19.69","Ages.70.90" )
    # Select data only for polygon of interest
    # converts to normal dataframe
    st_geometry(mapdata) <- NULL 
    # sum basic greenspace (where no options = TRUE)
    area.sums<-colSums(mapdata[,c(6:16)],na.rm=TRUE) # array

    # Calc chart data according to definition of greenspace
    chartdata<-data.frame(type=c("Allotments", "Golf Courses", "Open Access Land", "Parks & Gardens", "Playing fields",  "Religious & Cemetery", "Sports" ), area=c(area.sums[1],area.sums[4],area.sums[11],area.sums[8], sum(area.sums[c(6:7)]), sum(area.sums[c(3,9)]),sum(area.sums[c(2,5,10)])  ) ) 
    
    if(!alot) chartdata<-chartdata[-which(chartdata$type=="Allotments"),]
    if(!relg) chartdata<-chartdata[-which(chartdata$type=="Religious & Cemetery"),]
    if(!pkgd) chartdata<-chartdata[-which(chartdata$type=="Parks & Gardens"),]
    if(!golf) chartdata<-chartdata[-which(chartdata$type=="Golf Courses"),]
    if(!sports) chartdata<-chartdata[-which(chartdata$type=="Sports"),]
    if(!play) chartdata<-chartdata[-which(chartdata$type=="Playing fields"),] 
    if(!open) chartdata<-chartdata[-which(chartdata$type=="Open Access Land"),]
    
    chartdata[,"area"]<-chartdata[,"area"]/10000
    return(chartdata)
}


# Calculate bar chart data for chosen polygon on basis of selected greenspace map
calc_chartdata_2<-function(alot,relg,pkgd,golf,sports,play,open,mapdata,polygon){
    names(mapdata)<-c("soa","objectid","lsoa11nm","lsoa11nmw","st_lengths",   
                      "Allotments","Bowling_Green","Cemetery","Golf_Course","Sports" ,          
                      "Play_Space", "Playing_Field","Park_Garden","Religious_Grounds","Tennis_Court",     
                      "Open_Access", "All.Ages","Ages.0.18","Ages.19.69","Ages.70.90" )
    
    # Select data only for polygon of interest
    #print(class(mapdata))
    # converts to normal dataframe
    st_geometry(mapdata) <- NULL 
    mapdata<-mapdata[polygon,]
    # sum basic greenspace (where no options = TRUE)
    area.sums<-colSums(mapdata[,c(6:16)],na.rm=TRUE) # array

    # Calc chart data according to definition of greenspace
    chartdata<-data.frame(type=c("Allotments", "Golf Courses", "Open Access Land", "Parks & Gardens", "Playing fields",  "Religious & Cemetery", "Sports" ), area=c(area.sums[1],area.sums[4],area.sums[11],area.sums[8], sum(area.sums[c(6:7)]), sum(area.sums[c(3,9)]),sum(area.sums[c(2,5,10)])  ) ) 
    
    if(!alot) chartdata<-chartdata[-which(chartdata$type=="Allotments"),]
    if(!relg) chartdata<-chartdata[-which(chartdata$type=="Religious & Cemetery"),]
    if(!pkgd) chartdata<-chartdata[-which(chartdata$type=="Parks & Gardens"),]
    if(!golf) chartdata<-chartdata[-which(chartdata$type=="Golf Courses"),]
    if(!sports) chartdata<-chartdata[-which(chartdata$type=="Sports"),]
    if(!play) chartdata<-chartdata[-which(chartdata$type=="Playing fields"),] 
    if(!open) chartdata<-chartdata[-which(chartdata$type=="Open Access Land"),]
    
    chartdata[,"area"]<-chartdata[,"area"]/10000
    return(chartdata)
}

# Determine which polygon coorinates fall within
# https://github.com/r-spatial/sf/wiki/migrating

find_polygon <- function(polygons, point) {
  result<-st_intersects(polygons,point)
  final<-which(result==1)
  #polygon.sf<-polygons[final,]
  return(final)
}

```


```{r,context='data', include=FALSE}
# Load greenspace data 
greenspace.sf<-st_read(paste(dir_shape,"GrnSpc_by_SOA_map.shp",sep=""))

```
Greenspace Explorer
============================================================================================
Column {.sidebar}
-----------------------------------------------------------------------
### Define Greenspace

Greenspace includes allotments and parks & gardens, but might also be defined to include large areas such as golf courses and open access land.

Tick boxes to INCLUDE features:
```{r}

   checkboxInput("alot", "Include Allotments", value = TRUE)
   checkboxInput("relg", "Include Religious grounds", value = TRUE)
   checkboxInput("pkgd", "Include Parks & Gardens", value = TRUE)
   checkboxInput("play", "Include Playing fields & spaces", value = TRUE)
   checkboxInput("golf", "Include Golf courses", value = FALSE)
   checkboxInput("sports", "Include Sports facilities", value = TRUE)
   checkboxInput("open", "Include Open Access Areas", value = FALSE)
   
```

Note: maps recalculates after EVERY change - please be  patient!

### Map controls
```{r}
sliderInput("visibility.gsp","Map transparency:",min=0,max=1,value=0.7)

```

Column {data-width=600}
------------------------------------------------------------------------------------------------------------
### Amount of Greenspace by Population 
```{r}
leafletOutput("map")
```

```{r, context="server", include=FALSE}

#use reactive values to store the data you generate from a shape click
clickval <- reactiveValues()
clickval$polygon <- NULL
#clickval$soa<-NULL
  
# Reactive function to calculate mapdata and pallete for display returns both as list
map_change <- reactive({
  req(!is.null(input$alot))
  req(!is.null(input$relg))
  req(!is.null(input$pkgd))
  req(!is.null(input$golf))
  req(!is.null(input$sports))
  req(!is.null(input$play))
  req(!is.null(input$open))
  mapdata<-calc_grnsp(input$alot,input$relg,input$pkgd,input$golf,input$sports,input$play,input$open,
                      greenspace.sf) 
  
  q<-quantile(mapdata$area_per_pop,c(0,0.05,0.1,0.2,0.4,0.6,0.8,0.9,0.95,1),names=FALSE,na.rm=TRUE)
  bins <- unique(q)
  pal <- colorBin("Greens", domain = mapdata$area_per_pop, bins = bins, pretty=TRUE)
  fillcolour<-~pal(mapdata$area_per_pop)
  
  return(list(mapdata,fillcolour))
})
 

# Charts greenspace mix for selected polygon
chartdata_1<-reactive({
   req(!is.null(input$alot))
  req(!is.null(input$relg))
  req(!is.null(input$pkgd))
  req(!is.null(input$golf))
  req(!is.null(input$sports))
  req(!is.null(input$play))
  req(!is.null(input$open))
  # print(paste(input$alot,input$relg,input$pkgd,input$golf,input$sports,input$open,input$play))
  calc_chartdata_1( input$alot,input$relg,input$pkgd,input$golf,input$sports,input$play,input$open,greenspace.sf)
})

# Charts greenspace mix for selected polygon
chartdata_2<-reactive({
   req(!is.null(input$alot))
  req(!is.null(input$relg))
  req(!is.null(input$pkgd))
  req(!is.null(input$golf))
  req(!is.null(input$sports))
  req(!is.null(input$play))
  req(!is.null(input$open))
  req(!is.null(clickval$polygon))  
  #print(paste(input$alot,input$relg,input$pkgd,input$golf,input$sports,input$open,input$play,clickval$polygon))
  calc_chartdata_2( input$alot,input$relg,input$pkgd,input$golf,input$sports,input$play,input$open,greenspace.sf,clickval$polygon )
})

# Display basic map - take input defaults as first map - check
output$map <- renderLeaflet({
      leaflet() %>% 
      setView(lng = -4.8, lat = 50.5, zoom = 9) %>% 
      addTiles()  %>%  
      addScaleBar()
})

# Observe to change map Polygons if required
observe({
  # Define colours 
  visibility<-input$visibility.gsp
  parameters<-map_change()
  mapdata<-parameters[[1]]  
  fillcolour<-parameters[[2]]
  
  # Add information by SOA
  labels <- sprintf(
    "<strong>%s Ha of Greenspace for %g population</strong>",
    round(mapdata$area/10000,0), mapdata$All.Ages
  ) %>% lapply(htmltools::HTML)

  leafletProxy("map") %>%
   clearGroup("all") %>% 
    addPolygons(data=mapdata, group="all",
    fillColor = fillcolour,
    weight = 1,
    opacity = 1,
    color = "white",
    dashArray = "3",
    fillOpacity = visibility,
    highlight = highlightOptions(
      weight = 5,
      color = "#666",
      dashArray = "",
      fillOpacity = 0.7,
      bringToFront = TRUE),
      label = labels,
      labelOptions = labelOptions(
        style = list("font-weight" = "normal", padding = "3px 8px"),
        textsize = "15px",
        direction = "auto") ) %>%
    clearGroup("chosen") %>%  # add selected polygon
    addPolygons(data=greenspace.sf[isolate(clickval$polygon),], # use isolate render non-reactive to click
      group="chosen",
      fillColor = "red",
      weight = 1,
      opacity = 1,
      color = "white",
      dashArray = "3",
      fillOpacity = visibility )
  
})  

# Observe for click on polygon
# ref: https://stackoverflow.com/questions/41710104/r-plot-matching-selected-polygon-in-leaflet-map

observeEvent(input$map_shape_click, { # update the location selectInput on map clicks
  p <- input$map_shape_click
  point.sf<-st_point(c(p$lng,p$lat))
  # print(point.sf)
  clickval$polygon<-find_polygon(greenspace.sf,point.sf)
  #print(paste("clickval$polygon=",clickval$polygon))
  # OUTPUTS BASED ON CLICK VALUE
  #  output$coord_output <- renderText({
  #      paste("Polygon clicked: ",clickval$polygon, ". Population=",greenspace.sf$All_Ags[clickval$polygon])
  #  })

  # Create chart for selected polygon
  output$chart_soa <- renderPlot({
      ggplot(data = chartdata_2(), aes(x = type, y=area, fill=type)) +
      geom_bar(stat = "identity")  +
      labs(caption=paste("Selected polygon population=",greenspace.sf$All_Ags[clickval$polygon]),
           x="Types of Greenspace", y = "Area within SOA (hectares)", fill="")+
      theme_classic() +
      theme(axis.text.x=element_blank(),legend.position="right") 
  })
  
  # Outline selected polygon in different colour
  visibility<-input$visibility.gsp
  parameters<-map_change()
  mapdata<-parameters[[1]]
  fillcolour<-parameters[[2]]
  
   leafletProxy("map") %>%
    clearGroup("chosen") %>% 
    addPolygons(data=greenspace.sf[clickval$polygon,],
      group="chosen",
      fillColor = "red",
      weight = 1,
      opacity = 1,
      color = "white",
      dashArray = "3",
      fillOpacity = visibility )
    
})

```

Column {data-width=400}
------------------------------------------------------------------------------------------------------------
###  Greenspace composition: all Cornwall
```{r}
plotOutput("chart_all")
#textOutput("message_all")
```

###  Local composition: CLICK on map to select area

```{r}
plotOutput("chart_soa")
```

```{r, context='server'}
# Uses fillCol to make sizing responsice - alternative is simpluy to use render statements or to renderPlotly
# Use output in some way

output$chart_all <-   renderPlot({
      ggplot(data = chartdata_1(), aes(x = type, y=area, fill=type)) +
      geom_bar(stat = "identity")  +
      labs(x="Types of Greenspace", y = "Area in Cornwall (hectares)", fill="")+
      theme_classic() +
      theme(axis.text.x=element_blank(),legend.position="right") 
    })

fillCol(height = 600, flex = c(1, NA) , 
  #plotOutput("chart_all", height="45%", width="100%"),
  plotOutput("chart_soa", height="100%", width="100%"),
  textOutput("message")
)

output$diagram<-renderDiagrammeR({
DiagrammeR::mermaid("
graph TD
Z{OS Greenspaces}-->B[Trim to Cornwall & check projections]; Y{Open Access Land}-->B; 
B-->C[Merge Greenspace & Open access polygons]; 
C-->D[Greenspace area by type & SOA]; 
D-->E[Merge Greenspace and Population data by SOA]; 
X{Census Population by SOA}-->E
E-->F[Reproject and save for App display];
F-->G((Greenspace by Population))  "
, height=200)
})
```

Data Sources & Comments
============================================================================================
Column {data-width=400}
-------------------------------------
### Data Diagram{data-height=150 data-padding=20}
```{r}
DiagrammeROutput("diagram")
```

Column {data-width=600}
-------------------------------------
### Comments
Map uses data from:
(i) OS Greenspace, 
(ii) Natural England Open Access Land and 
(iii) Census population data.

The maps show the area of greenspace by population using "super output areas"" used by the Census and each representing a similar size of population (hence urban SOAs much smaller than rural SOAs). 

Including open access land has the largest effect on the distribution of greenspace per population, priviledging certain rural areas such as Bodmin moor.

Further Developments

1. Relatively simple additions might include:

  (i) Alternative polygons to classify greenspace - instead of SOAs one could use postcode areas, Neighbourhood planning areas, AONB areas or other groupings of interest. Fairly reliable estimates of population can usually be made for all but the smaller areas.
  (ii) Express greenspace as total greenspace area or as % of the total SOA area rather per head of population - the latter providing an indication of the spatial density of greenspace.
  (iii) Include additional types or measures of greenspace, or alternative measures such as the length of footpaths within SOA.

2. A similar approach could characterise areas by types of habitat or landcover or use, or indeed the density of traffic, noise etc per SOA/Postcode etc.

3. An alternative approach might calculate a 'greenspace indicator' based not only on the total area of greenspace but applying weights according to the % of the population that makes use of that type of space and has access to the area.

4. A complementary approach would provide information on access to greenspaces defined as the population within a set distance or travel time to a greenspace. From this a map of accessible greenspace per head of population might be developed (gridcell rather than polygon based) or a variety of other methods for displaying the information could be developed.